Keep your Laziness in Check

### 2018-03-08 paper/main.tex
[paper] wrote a subsection on connection between Demand a based specs and fromDemand-based specs

\section{Demand Types and Specification}
\label{sec:demand-types-and-spec}
There is a general pattern between a type \lk{a} and
its \textit{demand type}, for which we will write \lk{Demand a} from
now on. In both the example on lists and the example on maps,
a \lk{Thunk} type constructor is placed around all of the constructor
arguments to the original type. The \lk{Thunk} type has the following
definition:
\begin{inlinecode}
data Thunk a = E !a | T
\end{inlinecode}
Its two constructors \lk{E} and \lk{T} represents an evaluated
fragment of the original data, and unexplored regions of the original
data, respectively.
%
This design is motivated by the runtime representation of Haskell
values. Boxed values in Haskell are indirectly represented by a
pointer, and the pointer can lead to either a constructor bundled with
pointers to deeper parts of the value, or a piece of unevaluated code,
which will produce a fragment of the value upon execution
~\cite{SPJSTG}. This piece of code is also know as a thunk.

@@

>>>>

\section{Demand Types and Specification}
\label{sec:demand-types-and-spec}

\subsection{Types}
There is a general pattern between a type \lk{a} and
its \textit{demand type}, for which we will write \lk{Demand a} from
now on. In both the example on lists and the example on maps,
a \lk{Thunk} type constructor is placed around all of the constructor
arguments to the original type. The \lk{Thunk} type has the following
definition:
\begin{inlinecode}
data Thunk a = E !a | T
\end{inlinecode}
Its two constructors \lk{E} and \lk{T} represents an evaluated
fragment of the original data, and unexplored regions of the original
data, respectively.
%
This design is motivated by the runtime representation of Haskell
values. Boxed values in Haskell are indirectly represented by a
pointer, and the pointer can lead to either a constructor bundled with
pointers to deeper parts of the value, or a piece of unevaluated code,
which will produce a fragment of the value upon execution
~\cite{SPJSTG}. This piece of code is also know as a thunk.

@@

\subsection{Bridging ad-hoc laziness testing with StrictCheck}
In Haskell development, programmers often inserts divergent terms such
as \lk{error} or \lk{undefined} into the inputs of a function to test
its laziness. If a function evaluates such a divergent term, execution
gets aborted. We will call these divergent terms ``bottoms'' following
the common terminology in literature.

This is a very convenient approach to laziness testing because
programmers can use functions on these partial data containing bottoms
as if they are total values. As long as such functions don't require
the bottom values at runtime, execution continues. However, this
process typically involves repeated manual process of --- inserting
bottoms, executing the function, and diagnosing.

StrictCheck bridges the gap between this approach of laziness testing
and property-based laziness testing by supplying the two
functions \lk{fromDemand :: Demand a -> a} and \lk{toDemand :: a ->
Demand a}. Intuitively, \lk{fromDemand} produces a partial data
containing bottoms given a value of its demand type. This is performed
by re-assembing a value of type \lk{a} using the shaped described by
its demand value, and using a special bottom value at all locations
where a \lk{Thunk} appears in its demand value.

We implemented this process with a special exception, which is a kind
of bottom value, that cannot be thrown nor caught by users of
StrictCheck since its type is hidden. This allows \lk{toDemand} to
determine the precise locations of such bottoms, and convert them back
to the \lk{Thunk} value of type \lk{Demand a}.

In fact, in all of the 3 examples presented, we have written
specifications in this style instead of explicit manipulations of
type \lk{Demand a}. This gives us the full arsenal of existing API on
data types involved in the specification, and when runtime safety is
required, we simply convert these partial values back to demand
values.

Since StrictCheck can generate random demands on values, programmers
can eliminate the manual process inserting bottoms, and focus on the
diagnosis phase of the usual debug loop.
